### 2020-01-16 ISSTA/experiments.tex
big pass

In order to evaluate our approach and implementation, we applied
nondeterminism detection to real Python code, including some widely
used libraries (according to GitHub, {\tt pyfakefs} is used by at least
338 other projects, and {\tt redis-py} is used by at least 50,000).  The primary
points we wanted to explore were {\bf RQ1:} whether our approach was able to reliably detect actual
nondeterminism, {\bf RQ2:} whether the overhead of nondeterminism
detection for real systems was acceptable, and {\bf RQ3:} whether the performance of
delta-debugging in this setting was acceptable, in terms of both time
and results (amount of reduction in test size).  We chose two of
the subjects, {\tt redis-py} and {\tt pyfakefs}, based on the fact that they were previously-existing large
examples of TSTL harnesses, and therefore serve to show how easily our
approach can be integrated into an existing large-scale test
generation effort.  The other two subjects were chosen, and new harnesses
written, specifically to demonstrate specific, important sources of
nondeterminism in Python code: concurrency and hash salting.  In
addition to our core research questions ({\bf RQ1-RQ3}) shared by all subjects, we
report on particular aspects of using the TSTL nondeterminism tools
specific to each subject.

@@

There are several primary threats to validity:  first, our empirical
results are limited to a small set of Python programs, ranging from
relatively small and simple to large and complex libraries; the
representative nature of these subjects is not clear.
Furthermore, because
no other tools implement the kind of approach taken here, based on
automated generation of value pool based unit tests for a system
(imperative property-based testing), to our knowledge, we were unable
to perform a meaningful comparison with another nondeterminism
detection tool.  Available tools, such as DeFlaker \cite{bell2018d}
(\url{http://www.deflaker.org/}), perform a completely different task (DeFlaker
is essentially a plugin for Maven regression tests, and cannot check
nondeterminism at any smaller granularity than that of a whole tests's
pass/fail result) and target the Java language.  iDFlakies
\cite{idflakies}, similarly, ``does not further classify the causes of
the flaky tests'' beyond identifying them as likely related to test
order, or not.  However, the primary
aim of our results is to show 1) that nondeterminism at the library
level can exist in real Python programs, 2) that it can be detected
by an implementation of the formalism proposed in this paper, and 3)
that the overhead for such detection, and cost to delta-debug
nondeterministic tests, is not prohibitive.
\begin{comment}
We believe that the case
studies show that users testing Python libraries, and concerned about
nondeterminism, would find the TSTL tools useful, and would be able to
detect and debug nondeterministic behavior using our approach, a
functionality that we believe to be unique in the automated test generation
literature (and supported in no other tool).  Comparison with a
competing tool is not required to validate the basic idea, since
precise empirical measurements of overheads, or improvements on an
existing approach, are not proposed for evaluation.
\end{comment}
Another threat to validity is posed by implementation errors.
TSTL and the mutation tool are tested for bugs by a set of Travis CI
tests, and the nondeterminism
detected by TSTL can be demonstrated with standalone tests that do not depend on any TSTL code.
The source code for the tools and all experimental subjects is
available for inspection and replication in the TSTL repository.

>>>>

In order to evaluate our approach and implementation, we applied
nondeterminism detection to real Python code, including some widely
used libraries (according to GitHub, {\tt pyfakefs} is used by at least
338 other projects, and {\tt redis-py} is used by at least 50,000).  The primary
points we wanted to explore were {\bf RQ1:} whether our approach was able to reliably detect actual
nondeterminism, {\bf RQ2:} whether the overhead of nondeterminism
detection for real systems was acceptable, and {\bf RQ3:} whether the performance of
delta-debugging in this setting was acceptable, in terms of both time
and results (amount of reduction).  We chose two of
the subjects, {\tt redis-py} and {\tt pyfakefs}, based on the fact that they were previously-existing large
TSTL harnesses, and serve to show how easily our
approach can be integrated into an existing test
generation effort.  The other two subjects were chosen, and new harnesses
written, specifically to demonstrate specific, important sources of
nondeterminism in Python code: concurrency and hash salting.  In
addition to our core research questions ({\bf RQ1-RQ3}) shared by all subjects, we
report on aspects of using the TSTL nondeterminism tools
specific to each subject.

@@

There are several primary threats to validity:  first, our empirical
results are limited to a small set of Python programs, ranging from
relatively small and simple to large and complex libraries; the
representative nature of these subjects is not clear.
Furthermore, because
no other tools implement the kind of approach taken here, based on
automated generation of value pool based unit tests for a system
(imperative property-based testing), to our knowledge, we were unable
to perform a meaningful comparison with another nondeterminism
detection tool.  Available tools, such as DeFlaker \cite{bell2018d}
(\url{http://www.deflaker.org/}), perform a completely different task (DeFlaker
is essentially a plugin for Maven regression tests, and cannot check
nondeterminism at any smaller granularity than that of a whole tests's
pass/fail result) and target the Java language.  iDFlakies
\cite{idflakies}, similarly, ``does not further classify the causes of
the flaky tests'' beyond identifying them as likely related to test
order, or not.  However, the primary
aim of our results is to show 1) that nondeterminism at the library
level can exist in real Python programs, 2) that it can be detected
by an implementation of the formalism proposed in this paper, and 3)
that the overhead for such detection, and cost to delta-debug
nondeterministic tests, is not prohibitive.
\begin{comment}
We believe that the case
studies show that users testing Python libraries, and concerned about
nondeterminism, would find the TSTL tools useful, and would be able to
detect and debug nondeterministic behavior using our approach, a
functionality that we believe to be unique in the automated test generation
literature (and supported in no other tool).  Comparison with a
competing tool is not required to validate the basic idea, since
precise empirical measurements of overheads, or improvements on an
existing approach, are not proposed for evaluation.
\end{comment}
Another threat to validity is posed by implementation errors.
TSTL and the mutation tool are tested for bugs by a set of Travis CI
tests, and the nondeterminism
detected by TSTL can be demonstrated with standalone tests that do not
depend on our code.
The source code for our TSTL version and all experimental subjects will be
made available for inspection and replication.

### 2020-01-16 ISSTA/implementation.tex
big pass

While the formal definitions of nondeterminism proposed in Section
\ref{sec:formal} offer a framework for checking for nondeterministic
behavior, they leave unspecified the key notions of \emph{state},
\emph{visible state}, and \emph{actions}.  For example, we could
identify the state and visible state of a system as identical with the
full set of memory locations accessed by that system, and actions with the
individual processor instructions executed.  However, this
definition is highly impractical for two reasons.  First, it is
extremely inefficient to compare multiple executions for
full-memory-state equivalence at every instruction step.  Second, and
even more importantly, as noted in the introduction, very few programs
will behave deterministically when considered at this low level,
unless the full state of the operating system, processor, hardware
devices, etc. are all taken as controlled inputs.

What is a practical level of granularity for (visible) states and actions that
can be efficiently checked, and matches (for the most part) developer
and tester intuitions about what should behave deterministically?

Andrews et al. formalized a value-pool-based model of unit tests that is widely used in
actual testing tools, including Randoop \cite{Pacheco}, showing that
all Java unit tests (and, more generally, unit tests in similar
languages) can be represented in a \emph{canonical} form, where
a test consists of three parts:

\begin{enumerate}
\item A declaration of a set of array variables, referred to as
  \emph{value pools}, e.g., {\tt int [] intVP = new int[3];}.
  \item A set of assignments of constant values to elements of
    primitive type value pools, e.g. {\tt intVP[2] = 1;}.
    \item A part in which all statements are assignments of calls of a
      method to a value pool, with all arguments also
      taken from a pool, e.g. {\tt intVP[2] = fooVP[2].bar(intVP[0],
        intVP[1]);} (these are referred to as \emph{array-canonical} statements).
    \end{enumerate}

@@

We implemented our approach in the TSTL \cite{NFM15,ISSTA15,tstlsttt}
system, an open-source
language and tool for property-based testing
\cite{Hypothesis,ClaessenH00} of Python code.   TSTL, including full
support for our nondeterminism detection methods, is available at
\url{https://github.com/agroce/tstl}, and the {\tt examples} directory
contains the source code needed to duplicate all of our
experiments\footnote{In some cases, you will also need to install an
  appropriate version of the tested software, by doing, e.g., {\tt pip
    install pyfakefs}.}.
\begin{comment}
TSTL can also be installed using Python's {\tt pip} tool, by typing
{\tt pip install tstl}; you will need to clone the github repository
to have access to the case studies used in the paper, however.

@@

\subsection{Horizontal Nondeterminism Detection}
\label{sec:horizimp}

@@

Note that TSTL provides explicit support for exploring the real-world sources of
nondeterminism described in Section \ref{sec:sources}, by supporting
execution in a new process and allowing
control over timing changes.  Changing the timing of
steps is also effective in exposing concurrency-related
nondeterminism, since the delay will frequently change the
interleaving of threads.

Because TSTL has a first-class interface to AFL \cite{aflfuzz}, we can even use
AFL's sophisticated heuristics to perform very thorough, week-long
checks for nondeterminism, using strategies built to find subtle
security vulnerabilities in C programs.  Equally importantly, since
there is substantial overhead in nondeterminism detection, AFL can be
used to predict whether tested Python code is a good candidate for
checking for horizontal nondeterminism; if the
AFL \emph{stability} statistic (see the AFL documentation for details
\cite{aflfuzz}) is lower than 100\%, it may indicate horizontal
nondeterminism is present.
Because of the idiosyncrasies of AFL instrumentation and process
behavior, this is not always a reliable guide, but it is a very low
cost indicator produced as a by-product of fuzzing for other problems.

\subsection{Vertical Nondeterminism Detection}
\label{sec:vertimp}

First-class vertical nondeterminism checks are currently
limited to failures.  
%{\tt --checkFailureDeterminism} causes the TSTL
%compiler to emit a harness where every action that causes an \emph{expected}
%exception to be raised is repeated to ensure the same exception is
still raised.
%In terms of the formalism, if execution of action
$\alpha_i$ raises an exception that does not, itself, indicate a test failure, then 
    $F_i$ is true, and TSTL repeats $\alpha_i$ in order to ensure that
    $F_{i+1}$ is also true.
    %\end{enumerate}

>>>>

While the formal definitions above offer a framework for checking nondeterminism, they leave unspecified the key notions of \emph{state},
\emph{visible state}, and \emph{actions}.  For example, we could
identify the state and visible state of a system as identical with the
full set of memory locations accessed by that system, and actions with
processor instructions.  However, this
definition is highly impractical.  First, it is
extremely inefficient to compare multiple executions for
full-memory-state equivalence at every instruction step.  Second, and
even more importantly, as noted in the introduction, very few programs
are deterministic in this strict sense.  What is a practical level of granularity for (visible) states and actions that
can be efficiently checked, and matches (for the most part) developer
and tester intuitions about what should behave deterministically?

Andrews et al. formalized a value-pool-based model of unit tests that is widely used in
actual testing tools, including Randoop \cite{Pacheco}, showing that
unit tests can be represented in a \emph{canonical} form, where
a test consists of three parts:

\begin{enumerate}
\item A declaration of a set of array variables, referred to as
  \emph{value pools}, e.g., {\tt int [] intVP = new int[3];}.
  \item A set of assignments of constant values to elements of
    primitive type value pools, e.g. {\tt intVP[2] = 1;}.
    \item A part in which all statements are assignments of calls of a
      method to a value pool, with all arguments also
      taken from a pool, e.g. {\tt intVP[2] = fooVP[2].bar(intVP[0],
        intVP[1]);} (referred to as \emph{array-canonical} statements).
    \end{enumerate}

@@

We implemented our approach as a modification of the TSTL \cite{NFM15,ISSTA15,tstlsttt}
system, an open-source
language and tool for property-based testing
\cite{Hypothesis,ClaessenH00} of Python code.
%The TSTL {\tt examples} directory
%contains the SUTs needed to replicate our
%experiments.
\begin{comment}
TSTL can also be installed using Python's {\tt pip} tool, by typing
{\tt pip install tstl}; you will need to clone the github repository
to have access to the case studies used in the paper, however.

@@

\subsection{Nondeterminism Detection}
%\label{sec:horizimp}

@@

We provide support for the real-world sources of
nondeterminism described in Section \ref{sec:sources}, via
control over timing and/or execution in a new process.  Changing the timing of
steps is also effective in exposing concurrency-related
nondeterminism, since it often changes the
interleaving of threads.

Because TSTL has an interface to AFL \cite{aflfuzz}, we can use
AFL's sophisticated heuristics to perform very thorough, week-long
checks for nondeterminism.  Moreover, since
there is substantial overhead in nondeterminism detection, AFL can be
used to predict whether tested Python code is a good candidate for
checking for horizontal nondeterminism; if the
AFL \emph{stability} statistic (see the AFL documentation for details
\cite{aflfuzz}) is lower than 100\%, it may indicate horizontal
nondeterminism.
Because of the idiosyncrasies of AFL instrumentation and process
behavior, this is not always a reliable guide, but it is a very low
cost indicator produced as a by-product of fuzzing for other problems.

%\subsection{Vertical Nondeterminism Detection}
%\label{sec:vertimp}

First-class vertical nondeterminism checks are currently
limited to failures.  
%{\tt --checkFailureDeterminism} causes the TSTL
%compiler to emit a harness where every action that causes an \emph{expected}
%exception to be raised is repeated to ensure the same exception is
%still raised.
In terms of the formalism, if execution of action
$\alpha_i$ raises an exception that does not, itself, indicate a test failure, then 
    $F_i$ is true, and TSTL repeats $\alpha_i$ in order to ensure that
    $F_{i+1}$ is also true.
    %\end{enumerate}

### 2020-01-16 ISSTA/introduction.tex
big pass

Determinism is particularly important for testing and debugging, where
being able to exactly reproduce system behavior is essential to
productivity \cite{Firesmith}.  If executing ``the same test'' can produce
significantly different behavior each time it is run, the consequences
can be unfortunate.  Developers using a test exhibiting nondeterminism
to debug a system
face a challenge in reasoning about causality, in that an event
observed may not even take place the next time the test is run; this
pernicious phenomenon has been sometimes referred to as the dreaded
``Heisenbug'' \cite{Heisenbug}, though the more accurate usage is
``Mandelbug'' \cite{GrottkeBugs,FaultTriggers}, since Heisenbugs, properly speaking, are bugs that disappear or
alter behavior under instrumentation (especially in debugging).
Automated test generation systems using test coverage results to drive
the search for interesting inputs may go awry when a run executes
code only rarely.  And, most significantly, regression testing
effectiveness can be significantly reduced if tests fail only
intermittently and unpredictably, as a result of environmental
factors rather than bugs in changed code.  Such behavior is
unfortunately all too common:  Gao et
al. \cite{Gao:2015:MSU:2818754.2818764} observed coverage differences
of up to 184 lines of code for the same test, and false positive rates
as high as 96\%.  Nondeterminism is sometimes problematic for
developers, who will usually want to  assume that library code they
use behaves in a reliably predictable fashion; nondeterminism is
frequently vexing in debugging efforts; most importantly, perhaps, nondeterminism is \emph{often disastrous} for large-scale highly
automated testing.

@@

A system is deterministic if, given the system's complete state at a point in
time, it is possible, in principle, to predict its future behavior
perfectly.  We say ``in principle'' because in the real world,
prediction may be possible but hopelessly impractical.  We write
complex software systems in many cases because we cannot predict their
behavior (if we could perform the calculations in advance, ourselves,
we would just do so).  As a consequence, in software, rather than
defining determinism in terms of prediction, we usually 
say that a system is deterministic if, given the same state and
inputs, it always produces the same outputs.

Technically, many ``nondeterministic'' systems are not
nondeterministic in a strong sense, at all.  Given the complete state
of the system (which includes the entire state of the underlying
hardware, the operating system, storage devices, etc.),
most software \emph{is} completely predictable.  What we actually
mean is that, given a certain limited abstraction of state
and inputs, observable behavior is repeatable.  %This abstraction,
%for most software, is not expected to include many elements outside of
%the software system itself.

Because the
programmer's abstraction of the system ignores  such a large number of
details, very few tests are in fact ``deterministic'' in the sense
that they eliminate all changes in behavior between executions.  Running the same test
twice almost always results in differences, given a low enough level of
abstraction, since the system load,
cache contents, branch predictor history, etc. are almost never
controlled for; however, this kind of
nondeterminism is usually of no interest.  Rather, what matters is when
some kind of nondeterminism unexpectedly impacts computed values at a
higher level of abstraction, e.g. when it is believed that the
behavior of a thread scheduler will not matter, but a race condition
in the code means that it does matter, or when the order of items in
an iterator on a hash table is important, and the hash values used are
randomly salted.
\begin{comment} in general, if the operating system
itself is not buggy, low-level nondeterminism, by design, is invisible
except in fine-grained performance testing or real-time systems.
Unexpected nondeterminism usually arises when there is an element of higher-level
state or input that \emph{is} critical to the produced behavior, but
the programmer has not anticipated.  E.g.,  when it is believed that the
behavior of a thread scheduler will not matter, but a race condition
in the code means that it does matter, or when the order of items in
an iterator on a hash table is important, and the hash values used are
randomly salted.
\end{comment}

@@

Unexpected nondeterminism is, unfortunately, usually only discovered
in a context that makes it very hard to debug.   The most common such
contexts are occasional rare failures of a system in deployment, and
regression tests that do not behave reliably (known as flaky tests).
Furthermore, unexpected nondeterminism makes it difficult to use
automated test generation to produce effective regression tests for a
system.  While developers may know how to produce reliably
deterministic unit tests, automated test generation usually does not have sufficient information to avoid
producing the occasional such test.  Moreover, where a human might
make an assertion about the final value produced during a unit test,
an automatically generated regression test is usually most easily produced by
asserting equality with all values produced during a reference
run, since the final step of a test
is arbitrary, generated in an effort to increase
code coverage, not establish functional correctness.
This may be one reason that automated testing is seldom used to
produce general regression tests.  While failure-inducing tests
produced automatically are often added to regression suites, the full
set of passing tests is, to our knowledge, only infrequently preserved
as part of a typical regression suite, perhaps because such tests are
likely to be \emph{flaky} without considerable human effort.

@@

In order to help ensure that they are reliable and secure, complex
modern software systems usually include a large set of
\emph{regression tests}.  A regression test suite is a (usually large)
set of tests that can be run against a software system every time it
is modified, to ensure that the modification has not broken the system
in some way.  \emph{Flaky tests} \cite{miccoflaky} are regression
tests that fail in an intermittent, unreliable fashion, and thus
degrade the utility of regression testing.  The essence of a flaky
test is that, for the same snapshot of test code and code under test
(CUT), it sometimes fails and sometimes passes: the pass/fail result (disposition) of the
test is not a deterministic property of the test code, code under
test, and environment in which the test is run\footnote{Again, in 
  some sense the result \emph{is} clearly determined by the environment in
  which the test is run, but from the point of view of a developer or
  test engineer, the differences in environment are invisible and unknown.}.  This produces three serious
problems: first, a flaky test often wastes developer time and delays
software changes by forcing the investigation of code-under-test that
is not actually incorrect.  Second, failures in flaky tests (for that
reason) are often ignored, and therefore serious software faults may
be missed.  Finally, to mitigate these problems, flaky tests are often
run multiple times, wasting valuable computing resources and also
delaying acceptance of code changes.  An analogy may make the extent
of the problem in large-scale test automation clear:  a canary in a coal mine is of
little use if canaries frequently become ill for reasons unrelated to
the presence of toxic gases.  Mining may stop for no good reason, or
miners may learn to ignore the canary, leading to tragedy; a third,
more ``practical'' option is that miners may carry so many redundant
canaries into the coal mine that canary-care becomes a serious burden
on mining.

Flaky tests are, for us, simply a special case of general test
nondeterminism, where the nondeterminism of test values is sufficient
to cause the pass/fail result of the test to vary.  The definitions
and techniques proposed in this paper all apply to flaky tests, as a
special case of various kinds of \emph{horizontal} nondeterminism
described below.  Our focus, however, unlike all other tools and
approaches we are aware of, is on detecting the \emph{sources} of
flaky-ness \emph{before} they propagate to cause actual test failure.

@@

This paper proposes (1) a number of formal definitions of types of
nondeterminism (\emph{horizontal} and \emph{vertical}) (see Section \ref{sec:formal}) and (2) an implementation, based on these definitions, for detecting and debugging
nondeterminism in property-based testing.  The implementation is based
on an approach where (3) horizontal determinism is considered as a
kind of \emph{reflexive differential testing} (4) vertical
determinism is specialized to the common case of \emph{failure
  determinism}, and (5)  in both cases the formalism is made practical
by using the \emph{value pool} model of unit tests to define a useful
granularity for nondeterminism detection (see Section \ref{sec:pools}).
We further propose modifications to the widely used delta-debugging algorithm in order to better handle
nondeterminism as a test property.

>>>>

Determinism is particularly important for testing and debugging, where
being able to exactly reproduce system behavior is essential to
productivity \cite{Firesmith}.  If executing ``the same test'' can produce
significantly different behavior each time it is run, the consequences
can be unfortunate.  Developers using a test exhibiting nondeterminism
to debug a system
face a challenge in reasoning about causality, in that an event
observed may not even take place the next time the test is run; this
pernicious phenomenon has been sometimes referred to as the dreaded
``Heisenbug'' \cite{Heisenbug}, though the more accurate usage is
``Mandelbug'' \cite{GrottkeBugs,FaultTriggers}, since Heisenbugs, properly speaking, are bugs that disappear or
alter behavior under instrumentation (especially in debugging).
Automated test generation systems using test coverage results to drive
the search for interesting inputs may go awry when a run executes
code only rarely.  And, most significantly, regression testing
effectiveness can be significantly reduced if tests fail only
intermittently and unpredictably, as a result of environmental
factors rather than bugs in changed code.  Such behavior is
unfortunately all too common:  Gao et
al. \cite{Gao:2015:MSU:2818754.2818764} observed coverage differences
of up to 184 lines of code for the same test, and false positive rates
as high as 96\%.  Nondeterminism is often problematic for
developers, who want to  assume that library code behaves in a  predictable fashion; nondeterminism is
frequently vexing in debugging efforts; most importantly, perhaps, nondeterminism is \emph{often disastrous} for large-scale highly
automated testing.

@@

A system is deterministic if, given the system's complete state at a point in
time, it is possible, in principle, to predict its future behavior
perfectly.  We say ``in principle'' because in the real world,
prediction may be possible but  impractical.  We write
complex software systems because we cannot predict their
behavior (if we could perform the calculations in advance, 
we would just do so).  As a consequence, rather than
defining determinism in terms of prediction, we usually 
say that a system is deterministic if, given the same state and
inputs, it always produces the same outputs.  
Technically, many ``nondeterministic'' systems are not
nondeterministic in a strong sense.  Given the complete state
of the system (which includes the entire state of the underlying
hardware, the operating system, storage devices, etc.),
most software \emph{is} completely predictable.  What we actually
mean is that, \emph{given a certain limited abstraction of state
and inputs, observable behavior is repeatable}.  %This abstraction,
%for most software, is not expected to include many elements outside of
%the software system itself.

Because the
programmer's abstraction of the system ignores  such a large number of
details, very few tests are ``deterministic'' in the sense
that they eliminate all changes between executions.  Running the same test
twice almost always results in differences, given a low enough level of
abstraction, since the system load,
cache contents, branch predictor history, etc. are almost never
controlled for; however, this is seldom of interest.  What matters is when
some kind of nondeterminism unexpectedly impacts computed values at a
higher level of abstraction, e.g. when it is believed that the
behavior of a thread scheduler will not matter, but a race condition
in the code means that it does matter.

\begin{comment} in general, if the operating system
itself is not buggy, low-level nondeterminism, by design, is invisible
except in fine-grained performance testing or real-time systems.
Unexpected nondeterminism usually arises when there is an element of higher-level
state or input that \emph{is} critical to the produced behavior, but
the programmer has not anticipated.  E.g.,  when it is believed that the
behavior of a thread scheduler will not matter, but a race condition
in the code means that it does matter, or when the order of items in
an iterator on a hash table is important, and the hash values used are
randomly salted.
\end{comment}

@@

Unexpected nondeterminism is, unfortunately, usually only discovered
in a context that makes it very hard to debug.   The most common such
contexts are occasional rare failures of a system in deployment, and
regression tests that do not behave reliably (known as flaky tests).
Furthermore, unexpected nondeterminism makes it difficult to use
automated test generation to produce effective regression tests for a
system.  While developers may know how to produce reliably
deterministic unit tests, automated test generation usually does not have sufficient information to avoid
producing the occasional such test.  Moreover, where a human might
make an assertion about the final value produced during a unit test,
an automatically generated regression test is usually most easily produced by
asserting equality with all values produced during a reference
run, since the final step of a test
is arbitrary, generated in an effort to increase
code coverage, not establish functional correctness.
This may be one reason that automated testing is seldom used to
produce general regression tests.  While bugs found by tools are added
to regression suites, the full
set of passing tests is seldom preserved, perhaps because such tests are
likely to be \emph{flaky}.
%without considerable human effort.

@@

In order to help ensure that they are reliable and secure, complex
modern software systems usually include a large set of
\emph{regression tests}.  A regression test suite is a
set of tests that can be run against softwarre every time it
is modified, to ensure that the modification has not broken the system.  \emph{Flaky tests} \cite{miccoflaky} are regression
tests that fail in an intermittent, unreliable fashion.  The essence of a flaky
test is that, for the same snapshot of test code and code under test
(CUT), it sometimes fails and sometimes passes: the pass/fail result (disposition) of the
test is not a deterministic property of the test code, code under
test, and testing environment.  This produces three serious
problems: first, a flaky test often wastes developer time and delays
software changes by forcing the investigation of code-under-test that
is not actually incorrect.  Second, failures in flaky tests are often ignored, and therefore serious software faults may
be missed.  Finally, to mitigate these problems, flaky tests are often
run multiple times, wasting computing resources and
delaying acceptance of code changes.  As an analogy, we note that  a canary in a coal mine is of
little use if canaries frequently become ill for reasons unrelated to
the presence of toxic gases.  Mining may stop for no good reason, or
miners may learn to ignore the canary, leading to tragedy; a third,
more ``practical'' option is that miners may carry so many redundant
canaries into the coal mine that canary-care itself becomes a serious burden.

Flaky tests are, for us, simply a special case of general test
nondeterminism, where the nondeterminism of test values is sufficient
to cause the pass/fail result of the test to vary.  The definitions
and techniques proposed in this paper all apply to flaky tests, as a
special case of various kinds of \emph{horizontal} nondeterminism
described below.  Our focus, however, is on detecting sources of
flaky-ness in libraries, before they propagate.
%unlike all other tools and
%approaches we are aware of, is on detecting the \emph{sources} of
%flaky-ness \emph{before} they propagate to cause actual test failure.

@@

This paper proposes (1) a number of formal definitions of types of
nondeterminism (\emph{horizontal} and \emph{vertical}) and (2) an implementation, based on these definitions, for detecting and debugging
nondeterminism in property-based testing.  The implementation is based
on an approach where (3) horizontal determinism is considered as a
kind of \emph{reflexive differential testing} (4) vertical
determinism is specialized to the common case of \emph{failure
  determinism}, and (5)  in both cases the formalism is made practical
by using the \emph{value pool} model of unit tests.
We also introduce necessary modifications to the widely used delta-debugging algorithm in order to better handle
nondeterminism as a test property.

### 2020-01-16 ISSTA/relatedwork.tex
big pass

Bell et al. proposed DeFlaker \cite{bell2018d}, which makes flaky tests much easier to detect by relying on the observation that if a test fails, and does not cover any changed code then (1) presumably it was a passing test in the past, or the tests would not be ``green'' before the change so (2) the test is likely flaky.  Comparing our approach with DeFlaker is difficult; DeFlaker applies only in the context of code \emph{changes} and is essentially a heuristic that identifies failures as due to nondeterminism.  Our approach is primarily focused on automatically providing a more extensive \emph{specification} in automated test generation, where a test ``failing'' is itself a sign of nondeterministic behavior.  iDFlakies \cite{idflakies} is a framework and dataset for flaky tests, but again focuses on whole-tests, and detecting flakiness by actually observing it.  Their work shows the order-dependence accounts for a little more than 50\% of flaky test causes, and they (like us) classify the rest as due to concurrency, timeouts,
network/IO, etc.---the kinds of cause that our approach focuses on identifying.
The present paper, rather than focusing on flaky tests as such, investigates methods for handling nondeterminism in property-based test generation \cite{ClaessenH00,Papadakis:2011:PIT:2034654.2034663}.  From our point of view, flaky behavior is simply a special case of nondeterminism, where the nondeterminism is sufficient to cause the test to have different pass/fail results.  Our approach is, in a sense (especially the delta-debugging modifications) in line with Harman and O'Hearn's proposal to simply accept that ``All Tests Are Flaky'' \cite{StartupstoScaleups}, and work with probabilistically failing tests.  To our
knowledge, this is the first approach to the problem that
allows detection of potential sources of flakiness, based on real divergence in
behavior, without having to observe a test actually fail.  We additionally introduce the first variation of
delta-debugging that properly handles probabilistic reduction criteria.

>>>>

Bell et al. proposed DeFlaker \cite{bell2018d}, which makes flaky tests much easier to detect by relying on the observation that if a test fails, and does not cover any changed code then (1) presumably it was a passing test in the past, or the tests would not be ``green'' before the change so (2) the test is likely flaky.  Comparing our approach with DeFlaker is difficult; DeFlaker applies only in the context of code \emph{changes} and is essentially a heuristic that identifies failures as due to nondeterminism.  Our approach is primarily focused on automatically providing a more extensive \emph{specification} in automated test generation, where a test ``failing'' is itself a sign of nondeterministic behavior.  iDFlakies \cite{idflakies} is a framework and dataset for flaky tests, but again focuses on whole-tests, and detecting flakiness by actually observing it.  Their work shows the order-dependence accounts for a little more than 50\% of flaky test causes, and they (like us) classify the rest as due to ``concurrency, timeouts,
network/IO, etc.''---the kinds of cause that our approach focuses on identifying.
The present paper, rather than focusing on flaky tests as such, investigates methods for handling nondeterminism in property-based test generation \cite{ClaessenH00,Papadakis:2011:PIT:2034654.2034663}.  From our point of view, flaky behavior is simply a special case of nondeterminism, where the nondeterminism is sufficient to cause the test to have different pass/fail results.  Our approach is, in a sense (especially the delta-debugging modifications) in line with Harman and O'Hearn's proposal to simply accept that ``All Tests Are Flaky'' \cite{StartupstoScaleups}, and work with probabilistically failing tests.  To our
knowledge, this is the first approach to the problem that
allows detection of potential sources of flakiness, based on real divergence in
behavior, without having to observe a test actually fail.  We additionally introduce the first variation of
delta-debugging that properly handles probabilistic reduction criteria.
